Permutation and scalar transformation approaches to normalisation and combining them

### 2021-02-16 combining_perm-model.Rmd
spell check

```{r quadratic, fig.cap = "The raw wpd is plotted against nx*nfacet and the blue line represents the median of the multiple values for each nx*nfacet levels."}
G21 %>% 
  ggplot(aes(x=nx*nfacet, y = value)) +
  geom_point() + stat_summary(fun=median, geom="line", aes(group=1), color = "blue") + xlab("nx*nfacet") + ylab("raw wpd")

@@

A log-linear model is fitted to see how the values of $wpd_{raw}$ changes with the values of $nx$ and $nfacet$. The model is of the form $$y = a+b*log(x) +e$$, where $y = median(wpd_{raw})$ and $x = nx*nfacet$. $wpd_{lm}$ is a transformation on $wpd_{raw}$ which should be designed to remove the effect of $nx*nfacet$ on $wpd_{raw}$ and thus is defined as follows:
$wpd_{lm} = wpd_{raw} - a - b*log(nx*nfacet)$

```{r wpd-lm-horizontal,echo=FALSE, fig.caption = "$wpd_{lm}$ is platted again log(nx*nfacet) and this transformation leads to median($wpd_{lm}$) to having almost no relationship with log(nx*nfacet)"}

@@

In the earlier approach, $wpd_{raw} \in R$, whereas,  $0 \leq wpd_{raw}\leq 1$ since it is a JS distance. Also, JSD follows a Chi-square distribution, which could be considered as a Gamma distribution. Hence, to restrict the range of the measure between 0 and 1, and deviate from the normality assumption of the response variable in a OLS approach, we fit a generalized linear model (GLM) with the error distributed as a gamma distribution and link function as "inverse". Hence, the model is of the form  $$1/y = a+b*log(x) +e$$, where $y = median(wpd_{raw})$ and $x = nx*nfacet$. Again, $wpd_{glm}$ is a transformation on $wpd_{raw}$ which should be designed to remove the effect of $nx*nfacet$ on $wpd_{raw}$ and thus is defined as follows:
$wpd_{glm} = 1/wpd_{raw} - a - b*log(nx*nfacet)$.

Please note:

- subtracting the intercept brings the location of the transformed variable to 0 for either case
- heterogeneity is higher for this case after normalization as compared to the earlier one.

@@

 - Making the range of both the variables same by using min-max scaling method. In practice, however, we would only have one value of $wpd_{raw}$ which we need to transform using the modeling appraoch. Hence, min-max scaling approach could not be used here.

 - Standardizing the variables and expressing scores at standard deviation units. Again in practice, however, we would only have one value of $wpd_{raw}$ which we need to transform using the modeling appraoch. Hence, standardizing scores could not be used here as we do not have the mean and standard deviation of a series while using transfomation using modeling.

@@

The measure $wpd_{glm}$ seems to roughly follow a normal distribution except in the tails as could be seen in Figure \ref{fig:hist-qq} and the very method of permuation approach ensures that $wpd_permutation$ is also normally distributed. Further, they are brought to the similar scale and location and hence could be compared.

>>>>

```{r quadratic, fig.cap = "$wpd_{raw}$ is plotted against nx*nfacet and the blue line represents the median of the multiple values for each nx*nfacet levels."}
G21 %>% 
  ggplot(aes(x=nx*nfacet, y = value)) +
  geom_point() + stat_summary(fun=median, geom="line", aes(group=1), color = "blue") + xlab("nx*nfacet") + ylab("raw wpd")

@@

A log-linear model is fitted to see how the values of $wpd_{raw}$ changes with the values of $nx$ and $nfacet$. The model is of the form $$y = a+b*log(x) +e$$, where $y = median(wpd_{raw})$ and $x = nx*nfacet$. $wpd_{lm}$ is a transformation on $wpd_{raw}$ which should be designed to remove the effect of $nx*nfacet$ on $wpd_{raw}$ and thus is defined as follows:
$wpd_{lm} = wpd_{raw} - \hat a - \hat b*log(nx*nfacet)$

```{r wpd-lm-horizontal,echo=FALSE, fig.caption = "$wpd_{lm}$ is plotted again log(nx*nfacet) and this transformation leads to median($wpd_{lm}$) to having almost no relationship with log(nx*nfacet)"}

@@

In the earlier approach, $wpd_{raw} \in R$ was assumed, whereas,  $0 \leq wpd_{raw}\leq 1$ since it is a Jensen-Shannon Distance (JSD). Further, we know JSD follows a Chi-square distribution, which is a special case of Gamma distribution and hence belongs to exponential family of distributions. Therefore, we can fit a generalized linear model instead of a linear model to allow the response variable to follow a Gamma distribution. The inverse link is used when we know that the mean response to be bounded, which is applicable in our case since JSD will lie between between 0 and 1.

We fit the Gamma generalized linear model with inverse link which is of the form:  
$$y = 1/(a+b*log(x)) + e$$, where $y = median(wpd_{raw})$, $x = nx*nfacet$. Again, $wpd_{glm}$ which is a transformation on $wpd_{raw}$ and is designed to remove the effect of $nx*nfacet$ on $wpd_{raw}$ are chosen as the residuals and defined as
$wpd_{glm} = wpd_{raw} - 1/(\hat a + \hat b*log(nx*nfacet))$.

<!-- Please note: -->

<!-- - subtracting the intercept brings the location of the transformed variable to 0 for either case -->
<!-- - heterogeneity is higher for this case after normalization as compared to the earlier one. -->

@@

 - Making the range of both the variables same by using min-max scaling method. In practice, however, we would only have one value of $wpd_{raw}$ which we need to transform using the modeling approach. Hence, min-max scaling approach could not be used here.

 - Standardizing the variables and expressing scores at standard deviation units. Again in practice, however, we would only have one value of $wpd_{raw}$ which we need to transform using the modeling approach. Hence, standardizing scores could not be used here as we do not have the mean and standard deviation of a series while using transformation using modeling.

@@

The measure $wpd_{glm}$ seems to roughly follow a normal distribution except in the tails as could be seen in Figure \ref{fig:hist-qq} and the very method of permutation approach ensures that $wpd_permutation$ is also normally distributed. Further, they are brought to the similar scale and location and hence could be compared.
